---
output: github_document
bibliography: vignettes/chantrics.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# chantrics

<!-- badges: start -->
![R-CMD-check](https://github.com/tbruckbauer/chantrics/workflows/R-CMD-check/badge.svg?branch=master)
[![codecov](https://codecov.io/gh/tbruckbauer/chantrics/branch/master/graph/badge.svg?token=EQQ177WODM)](https://codecov.io/gh/tbruckbauer/chantrics)
![pkgdown](https://github.com/tbruckbauer/chantrics/workflows/pkgdown/badge.svg)
[![CodeFactor](https://www.codefactor.io/repository/github/tbruckbauer/chantrics/badge?s=8292aacb6a947f6280b202c4f29e21e4510dce53)](https://www.codefactor.io/repository/github/tbruckbauer/chantrics)
<!-- badges: end -->

`chantrics` applies the Chandler-Bate loglikelihood adjustment [@chanbate07] implemented in the [chandwich](https://cran.r-project.org/package=chandwich) package to different models frequently used in basic Econometrics applications. `adj_loglik()` adjusts the parameter covariance matrix of the models to incorporate clustered data, and can mitigate for model misspecification by wrapping `chandwich::adjust_loglik` for the supported models.

The returned model of class `chantrics` can be plugged into standard model evaluation and model comparison methods, for example `summary()`, `confint()` and `anova`, and a hypothesis test framework provided by `alrtest()`.

## Installation

<!--You can install the released version of chantrics from [CRAN](https://CRAN.R-project.org) with:

```{r}
# install.packages("chantrics")
```
-->

You can install the development version from [GitHub](https://github.com/) with:

```{r eval=FALSE}
## Remove the # if "devtools" is not installed yet.
# install.packages("devtools")
devtools::install_github("tbruckbauer/chantrics")
```
## Usage

This example is using the misspecified count data regression from Chapter 5.1 in the Object-Oriented Computation of Sandwich Estimators vignette from the *sandwich* package [@zeileis06]. First, data from a negative binomial model is generated, and then a Poisson model is fit, which is clearly misspecified.

```{r}
library(chantrics)

set.seed(123)
x <- rnorm(250)
y <- rnbinom(250, mu = exp(1 + x), size = 1)

## Fit the Poisson glm model, which is not correctly specified
fm_pois <- glm(y~x +I(x^2), family = poisson)
lmtest::coeftest(fm_pois)
# The I(x^2) term is spuriously significant.
```

In order to mitigate the misspecification, the loglikelihood adjustment is applied to the model object using `adj_loglik(model_object)`. If clustered data is available, a vector or factor indicating from which cluster the observation originates can be passed into the function using `cluster`. If it is not supplied, it is assumed that each observation originates from its own cluster and is independent. More information on this can be found in the [clustering vignette](https://chantrics.theobruckbauer.eu/articles/clustering-vignette.html).

```{r}
## Apply the loglikelihood adjustment to the model
fm_pois_adj <- adj_loglik(fm_pois)
lmtest::coeftest(fm_pois_adj)
# The I(x^2) term is no longer significant.
```

After this, some of the functionality is demonstrated. More information on the different methods can be found in the [Introducing chantrics vignette](https://chantrics.theobruckbauer.eu/articles/chantrics-vignette.html).

## References
