---
title: "An introduction to chantrics"
author: "Theo Bruckbauer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: chantrics.bib
vignette: >
  %\VignetteIndexEntry{An introduction to chantrics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r knitr_init, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(chantrics)
```

`chantrics` applies the Chandler-Bate loglikelihood adjustment [@chanbate07] implemented in the [chandwich](https://cran.r-project.org/package=chandwich) package [@chandwich] to different models frequently used in basic Econometrics applications. `adj_loglik()` is the central function of `chantrics`, it is a generic function adjusting the parameter covariance matrix of the models to incorporate clustered data, and can mitigate for model misspecification. The returned object can then be plugged into a range of model analysis functions, which will be described below. [For an introduction on modelling with clustered data using `chantrics`, visit this vignette.](https://chantrics.theobruckbauer.eu/articles/clustering-vignette.html).



## Functionality
*Note that not all functionality demonstrated below is available for all types of models.*

In order to be able to demonstrate the range of functionality available, this example will be using the misspecified count data regression from Chapter 5.1 in the Object-Oriented Computation of Sandwich Estimators vignette from the *sandwich* package [@zeileis06]. 

First, data from a negative binomial model is generated, and then a Poisson model is fit, which is clearly misspecified.

```{r datagen}
set.seed(123)
x <- rnorm(250)
y <- rnbinom(250, mu = exp(1 + x), size = 1)

## Fit the Poisson glm model, which is not correctly specified
fm_pois <- glm(y~x +I(x^2), family = poisson)
lmtest::coeftest(fm_pois)
# The I(x^2) term is spuriously significant.
```

We can now use the model object `fm_pois`, and adjust it using `adj_loglik()`. Use `coef()` to get a vector of the coefficients, `summary()` to get an overview over the adjustment, or use `lmtest::coeftest` to see the results of \(z\) tests on each of the coefficients.

```{r adjust_fmpois}
fm_pois_adj <- adj_loglik(fm_pois)
coef(fm_pois_adj) # class "numeric"
summary(fm_pois_adj)
lmtest::coeftest(fm_pois_adj)
```

### Confidence intervals of the estimates

The function `chandwich::conf_intervals()` returns confidence intervals at the level specified in `conf` (default: `95`). To use one of the other specifications of the adjustment from @chanbate07, use the `type` argument. Many other adjustments are available. The classic S3 method `confint()` is also available.

```{r ci}
chandwich::conf_intervals(fm_pois_adj)
chandwich::conf_intervals(fm_pois_adj, type = "spectral", conf = 99)
confint(fm_pois_adj)
```

We can also plot confidence regions of the estimates for two coefficients using `conf_region()`, where we can specify the parameters using `which_pars`, the type of specification of the adjustment from @chanbate07  using `type`, and the confidence levels using `conf`. Other adjustments are available.

```{r conf_region, fig.align='center', fig.width=7, fig.height=7}
fm_pois_adj_vert <- chandwich::conf_region(fm_pois_adj, which_pars = c("x", "I(x^2)"))
fm_pois_adj_none <- chandwich::conf_region(fm_pois_adj, which_pars = c("x", "I(x^2)"), type = "none")
plot(fm_pois_adj_vert, fm_pois_adj_none, conf = c(60, 80, 90, 95), col = c("brown", "darkgreen"), lty = c(1, 2), lwd = 2.5)
```

### Creating nested models

```{r update}
fm_pois_small_adj <- update(fm_pois_adj, .~.-I(x^2))

```

### Other diagnostic functions
The methods 
* `AIC()` for the Akaike Information Criterion, 
* `df.residual()` for the degrees of freedom of the residuals
* `fitted()` for the fitted values, 
* `logLik()` for the sum of the loglikelihoods and `logLik_vec()` for the loglikelihood contributions for each of the observations, and
* `vcov()` for the variance-covariance matrix,
are available.


The performance of the types of adjustments that are shown in @chanbate07 can be seen using `plot()` if there is a single free parameter. Use `type` to specify the types of adjustment that should show in the plot.

```{r plot_chan, fig.align='center', fig.width=7, fig.height=7}
fm_pois_smallest_adj <- update(fm_pois_small_adj, .~.-1)
plot(fm_pois_smallest_adj, type = 1:4, col = 1:4, legend_pos = "bottom", lwd = 2.5)
```

Note that for one free parameter, the Cholesky and the spectral adjustments are identical, and the vertical adjustment only deviates slightly at the edges of the plot.

## Supported models

